Tensor denoising and completion based on ordinal observations

### 2020-03-02 paper/arxiv/correction.tex
upload correction

To facilitate the examination of MSE, we define an order-$(K+1)$ tensor, $\tZ=\entry{z_{\omega,\ell}}\in\mathbb{R}^{d_1\times \cdots \times d_K\times (L-1)}$, by stacking the parameters $\Theta=\entry{\theta_\omega}$ and $\mb=\entry{b_\ell}$ together. Specifically, let $z_{\omega,\ell}=-\theta_\omega+b_\ell$ for all $\omega\in[d_1]\times \cdots \times [d_K]$ and $\ell\in[L-1]$; that is,
\[
\tZ=-\Theta\otimes \mathbf{1}+\tJ\otimes \mb,
\]
where $\mathbf{1}$ denotes a length-$(L-1)$ vector of all ones. Under the identifiability constraint $\langle \Theta, \tJ \rangle=0$, there is an one-to-one mapping between $\tZ$ and $(\Theta, \mb)$, with $\text{rank}(\tZ)=(\text{rank}(\Theta),2)^T$. Furthermore, 
\begin{equation}\label{eq:decomp}
\FnormSize{}{\hat \tZ-\trueZ}^2=\FnormSize{}{\hat \Theta-\trueT}^2 (L-1)+\FnormSize{}{\hat \mb-\trueb}^2 \left(\prod_kd_k\right),
\end{equation}
where $\trueZ=\trueT\otimes \mathbf{1}+\tJ\otimes \trueb$ and $\hat \tZ=\hat \Theta\otimes \mathbf{1}+\tJ\otimes \hat \mb$.

@@

We define the following constants:
\begin{align}\label{eq:constantZ}
C_{\alpha,\beta,\Delta}&=\max_{|z|\leq \alpha+\beta}\max_{\substack{z'\leq z-\Delta\\z''\geq z+\Delta}}\max\left\{ {\dot{f}(z) \over f(z)-f(z')},\ {\dot{f}(z) \over f(z'')-f(z)}\right\},\\
D_{\alpha,\beta,\Delta}&=\max_{|z|\leq \alpha+\beta}\max_{\substack{z'\leq z-\Delta\\z''\geq z+\Delta}}\max\left\{- {\partial\over \partial z}\left({\dot{f}(z) \over f(z)-f(z')}\right),\  {\partial\over \partial z}\left({\dot{f}(z) \over f(z'')-f(z)}\right) \right\},\\
A_{\alpha,\beta,\Delta}&=\min_{|z|\leq \alpha+\beta}\min_{z'\leq z-\Delta} \left(f(z)-f(z')\right).
\end{align}

@@

\begin{thm}[Statistical convergence with unknown $\mb$]\label{thm:ratejoint}
Consider an ordinal tensor $\tY\in[L]^{d_1\times \cdots \times d_K}$ generated from model~\eqref{eq:model} with the link function $f$ and parameters $(\trueT, \trueb)\in\tP\times \tB$. Suppose the link function $f$ satisfies Assumption~\ref{ass:joint}. Define $r_{\max}=\max_k r_k$, and assume $r_{\max}=\tO(1)$.

@@

Let $\mz_\omega=(z_{\omega,1},\ldots,z_{\omega,L-1})^T\in\mathbb{R}^{L-1}$ and $\mz=(\mz_{1,\ldots,1},\ldots,\mz_{d_1,\ldots,d_K})^T\in\mathbb{R}^{(L-1)\prod_k d_k}$. It follows from property (a) that
\[
\mz^T\mH\mz =\sum_{\omega}\mz_\omega^T \mH_{\omega} \mz_\omega.
\]
Furthermore, properties (b) and (c) (or similar arguments as in page 29, arXiv preprint) implies that
\begin{equation}
\mz_\omega^T \mH_{\omega} \mz_\omega \leq -D_{\alpha,\beta,\Delta} \sum_{\ell}z^2_{\omega,\ell}\KeepStyleUnderBrace{\mathds{1}_{\{y_\omega=\ell \text{ or }\ell+1\}}}_{\text{Bernoulli r.v.\ with probability bounded by $A_{\alpha,\beta,\Delta}$}}.
\end{equation}
Therefore,
\begin{equation}\label{eq:sum}
\mz^T\mH\mz =\sum_{\omega}\mz_\omega^T \mH_{\omega} \mz_\omega \leq -D_{\alpha,\beta,\Delta}\sum_{\omega}\sum_{\ell} z^2_{\omega,\ell}\mathds{1}_{\{y_\omega=\ell \text{ or }\ell+1\}}
\end{equation}
Based on central limit theorem, as the tensor dimension goes to infinity, 
\begin{equation}\label{eq:clt}
\sum_{\omega}\sum_{\ell} z^2_{\omega,\ell}\mathds{1}_{\{y_\omega=\ell \text{ or }\ell+1\}} \rightarrow\sum_{\omega}\sum_{\ell} z^2_{\omega,\ell}\mathbb{P}(y_\omega=\ell \text{ or }\ell+1)\geq c_2A_{\alpha,\beta,\Delta}\FnormSize{}{\mz}^2. 
\end{equation}
By~\eqref{eq:sum} and~\eqref{eq:clt}, we have
\begin{equation}\label{eq:boundH}
\mz^T\mH\mz \leq-c_2A_{\alpha,\beta,\Delta}D_{\alpha,\beta,\Delta}\FnormSize{}{\mz}^2,
\end{equation}
and therefore~\eqref{eq:HessionZ} is proved. Plugging~\eqref{eq:scoreZ} and~\eqref{eq:HessionZ} into~\eqref{eq:boundZ} yields
\[
\FnormSize{}{\hat \tZ-\trueZ}^2 \leq {c_1r^{K-1}_{\max}C^2_{\alpha,\beta,\Delta}\over A^2_{\alpha,\beta,\Delta}D^2_{\alpha,\beta,\Delta}}\left(L-1+\sum_k d_k\right).
\]
The MSEs for $\hat \Theta$ and $\hat \mb$ readily follow from~\eqref{eq:decomp}. 
\end{proof}

>>>>

To facilitate the examination of MSE, we define an order-$(K+1)$ tensor, $\tZ=\entry{z_{\omega,\ell}}\in\mathbb{R}^{d_1\times \cdots \times d_K\times (L-1)}$, by stacking the parameters $\Theta=\entry{\theta_\omega}$ and $\mb=\entry{b_\ell}$ together. Specifically, let $z_{\omega,\ell}=-\theta_\omega+b_\ell$ for all $\omega\in[d_1]\times \cdots \times [d_K]$ and $\ell\in[L-1]$; that is,
\[
\tZ=-\Theta\otimes \mathbf{1}+\tJ\otimes \mb,
\]
where $\mathbf{1}$ denotes a length-$(L-1)$ vector of all ones. Under the identifiability constraint $\langle \Theta, \tJ \rangle=0$, there is an one-to-one mapping between $\tZ$ and $(\Theta, \mb)$, with $\text{rank}(\tZ)\leq (\text{rank}(\Theta),2)^T$. Furthermore, 
\begin{equation}\label{eq:decomp}
\FnormSize{}{\hat \tZ-\trueZ}^2=\FnormSize{}{\hat \Theta-\trueT}^2 (L-1)+\FnormSize{}{\hat \mb-\trueb}^2 \left(\prod_kd_k\right),
\end{equation}
where $\trueZ=\trueT\otimes \mathbf{1}+\tJ\otimes \trueb$ and $\hat \tZ=\hat \Theta\otimes \mathbf{1}+\tJ\otimes \hat \mb$.

@@

We define the following constants that will be used in the theory:
\begin{align}\label{eq:constantZ}
C_{\alpha,\beta,\Delta}&=\max_{|z|\leq \alpha+\beta}\max_{\substack{z'\leq z-\Delta\\z''\geq z+\Delta}}\max\left\{ {\dot{f}(z) \over f(z)-f(z')},\ {\dot{f}(z) \over f(z'')-f(z)}\right\},\\
D_{\alpha,\beta,\Delta}&=\max_{|z|\leq \alpha+\beta}\max_{\substack{z'\leq z-\Delta\\z''\geq z+\Delta}}\max\left\{- {\partial\over \partial z}\left({\dot{f}(z) \over f(z)-f(z')}\right),\  {\partial\over \partial z}\left({\dot{f}(z) \over f(z'')-f(z)}\right) \right\},\\
A_{\alpha,\beta,\Delta}&=\min_{|z|\leq \alpha+\beta}\min_{z'\leq z-\Delta} \left(f(z)-f(z')\right).
\end{align}

@@

\begin{rmk} Add the specific bound for logistic link. 
\end{rmk}
\begin{thm}[Statistical convergence with unknown $\mb$]\label{thm:ratejoint}
Consider an ordinal tensor $\tY\in[L]^{d_1\times \cdots \times d_K}$ generated from model~\eqref{eq:model} with the link function $f$ and parameters $(\trueT, \trueb)\in\tP\times \tB$. Suppose the link function $f$ satisfies Assumption~\ref{ass:joint}. Define $r_{\max}=\max_k r_k$, and assume $r_{\max}=\tO(1)$.

@@

Let $\mz_\omega=(z_{\omega,1},\ldots,z_{\omega,L-1})^T\in\mathbb{R}^{L-1}$ and $\mz=(\mz_{1,\ldots,1},\ldots,\mz_{d_1,\ldots,d_K})^T\in\mathbb{R}^{(L-1)\prod_k d_k}$. It follows from property (a) that
\[
\mz^T\mH\mz =\sum_{\omega}\mz_\omega^T \mH_{\omega} \mz_\omega.
\]
Furthermore, properties (b) and (c) (or similar arguments as in page 29, arXiv preprint) imply that
\begin{equation}
\mz_\omega^T \mH_{\omega} \mz_\omega \leq -D_{\alpha,\beta,\Delta} \sum_{\ell}z^2_{\omega,\ell}\KeepStyleUnderBrace{\mathds{1}_{\{y_\omega=\ell \text{ or }\ell+1\}}}_{\text{Bernoulli r.v.\ with probability bounded by $A_{\alpha,\beta,\Delta}$}}.
\end{equation}
Therefore,
\begin{equation}\label{eq:sum}
\mz^T\mH\mz =\sum_{\omega}\mz_\omega^T \mH_{\omega} \mz_\omega \leq -D_{\alpha,\beta,\Delta}\sum_{\omega}\sum_{\ell} z^2_{\omega,\ell}\mathds{1}_{\{y_\omega=\ell \text{ or }\ell+1\}}.
\end{equation}
Based on central limit theorem, as the tensor dimension goes to infinity, 
\begin{equation}\label{eq:clt}
\sum_{\omega}\sum_{\ell} z^2_{\omega,\ell}\mathds{1}_{\{y_\omega=\ell \text{ or }\ell+1\}} \rightarrow\sum_{\omega}\sum_{\ell} z^2_{\omega,\ell}\mathbb{P}(y_\omega=\ell \text{ or }\ell+1)\geq c_2A_{\alpha,\beta,\Delta}\FnormSize{}{\mz}^2
\end{equation}
holds with very high probability. 
By~\eqref{eq:sum} and~\eqref{eq:clt}, we have
\begin{equation}\label{eq:boundH}
\mz^T\mH\mz \leq-c_2A_{\alpha,\beta,\Delta}D_{\alpha,\beta,\Delta}\FnormSize{}{\mz}^2,
\end{equation}
and therefore~\eqref{eq:HessionZ} is proved. Plugging~\eqref{eq:scoreZ} and~\eqref{eq:HessionZ} into~\eqref{eq:boundZ} yields
\[
\FnormSize{}{\hat \tZ-\trueZ}^2 \leq {c_1r^{K-1}_{\max}C^2_{\alpha,\beta,\Delta}\over A^2_{\alpha,\beta,\Delta}D^2_{\alpha,\beta,\Delta}}\left(L-1+\sum_k d_k\right).
\]
The MSEs for $\hat \Theta$ and $\hat \mb$ readily follow from~\eqref{eq:decomp}. 
\end{proof}
